---
title: "eds213-job-risk"
format: html
---

```{r, message=FALSE}
library(tidyverse)
library(DBI)
library(duckdb)
library(dplyr)
```

### What is the AI Automation Risk Across MEDS Job Postings Position Types? 
We will use the AI Penetration score average for all tasks associated with each position type. This score measures the degree to which Claude could complete a task. 

```{r}
# Connect to job risk databse
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir = "job_risk.duckdb")

# Create lazy Reference to database tables
# Tables do not exist yet, but can be accessed using SQL
meds_jobs_db      <- tbl(conn, "meds_jobs")
meds_jobs_tasks_db <- tbl(conn, "meds_jobs_tasks")
task_penetration_db <- tbl(conn, "task_penetration")
```

```{r}
# AI Penetration Score by Position Type
position_risk <- meds_jobs_tasks_db %>% 
  inner_join(task_penetration_db, by = "task") %>% # Use inner join to keep all ways for both tables
  inner_join(meds_jobs_db, by = "job_title") %>% # Join on job titles since they are distinct
  group_by(position_type) %>% 
  summarise(
    avg_ai_penetration = round(mean(penetration, na.rm = TRUE), 3),
    num_tasks = n()
  ) %>% 
  arrange(desc(avg_ai_penetration)) %>% 
  head(10) %>% 
  collect()

position_risk
```

### Visualize AI Risk for Position Types

```{r}
position_risk %>% 
  filter(!is.na(position_type)) %>% # Remove NA to easy plotting
  mutate(position_type = str_to_title(position_type)) %>% # Turn position type column to title to make letters official
  arrange(avg_ai_penetration) %>% 
  mutate(position_type = factor(position_type, levels = position_type)) %>% # Factor position type based on AI penetration score
  ggplot(aes(x = position_type, y = avg_ai_penetration)) +
  geom_segment(aes(xend = position_type, yend = 0), linewidth = 1.3, color = "darkgreen") +
  geom_point(size = 6, color = "darkgreen") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Average AI Task Penetration Score",
    title = "Which Position Types Are Most Exposed To AI?",
    subtitle = "Ranked by average AI task penetration score across MEDS job postings"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray40"),
    axis.text.y = element_text(size = 11)
  )
```

### What is the AI Automation Risk Across MEDS Job Postings Sectors? 
We will use the AI Penetration score average for all tasks associated with each sector. This score measures the degree to which Claude could complete a task. 

```{r}
# AI Penetration Score by Sector
sector_risk <- meds_jobs_tasks_db %>% 
  inner_join(task_penetration_db, by = "task") %>% # Use inner join to keep all ways for both tables
  inner_join(meds_jobs_db, by = "job_title") %>%  # Join on job titles since they are distinct
  group_by(sector) %>% 
  summarise(
    avg_ai_penetration = round(mean(penetration, na.rm = TRUE), 3),
    num_tasks = n()
  ) %>% 
  arrange(desc(avg_ai_penetration)) %>% 
  collect()

sector_risk
```

```{r}
# Disconnect from database to avoid permission issues in future
DBI::dbDisconnect(conn, shutdown = TRUE)
```

### Visualize AI Risk for Sector

```{r}
sector_risk %>% 
  arrange(avg_ai_penetration) %>% 
  mutate(sector = factor(sector, levels = sector)) %>% # Set order through levels by factoring
  ggplot(aes(x = sector, y = avg_ai_penetration)) +
  geom_segment(aes(xend = sector, yend = 0), linewidth = 1.3, color = "darkblue") +
  geom_point(size = 6, color = "darkblue") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Average AI Task Penetration Score",
    title = "Which Sectors Are Most Exposed To AI?",
    subtitle = "Ranked by average AI task penetration score across MEDS job postings"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray40"),
    axis.text.y = element_text(size = 11)
  )
```


